---
title: "Homework 1"
author: "Yiran Liu (netID: yiranl7)"
date: "1/29/2020"
output:
  pdf_document:
    latex_engine: xelatex
mainfont: Georgia
monofont: Sarasa Term CL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

To obtain the Monte Carlo estimation of $E(e^{X^2})$, we need to draw i.i.d. samples ${x_1, x_2, ..., x_n}$ from $U(0,1)$ and then calculate $\frac{1}{n}(e^{x_1}+e^{x_2}+...+e^{x_n})$. Codes and results:

```{r}
n <- 1000
x <- runif(n)
ex <- exp(x^2)
mu <- mean(ex)
se <- sd(ex)/sqrt(n)
```

The Monte Carlo estimate of $E(e^{X^2})$ is `r mu`, and the standard error of the estimate is `r se`.

## Question 2

To apply inversion method, we need to obtain the cdf of Pareto distribution first:
$$
F(x)=\int^{x}_{2}160t^{-6}dt=1-32x^{-5},\text{  }2\leq x <\infty
$$
And the inverse of $F(x)$ is
$$
F^{-1}(t)={(\frac{32}{1-t})}^{\frac{1}{5}}
$$
Then we draw i.i.d. samples ${u_1, u_2, ..., u_n}$ from $U(0,1)$ and plug in $F^{-1}(t)$. According to inversion method, the ${X_i=F^{-1}(u_i)}$ is the desired random sample of the Pareto distribution $f(x)$. Codes and results:

```{r}
n <- 1000
u <- runif(n)
x <- (32 / (1 - u))^(1/5)
```

The mean estimation is `r mean(x)` and the variance estimation is `r var(x)`.

## Question 3

As long as we have the inverse of $f(x)$, we can easily apply inversion method to get the random sample from $f(x)$. By definition, we have the cdf
$$
F(x) = 
\begin{cases}
0 & x<0 \\
\frac{1}{2}x^2 & 0\leq x \leq 1 \\
-\frac{1}{2}x^2+2x-1 & 1<x\leq2\\
1 & x>2 \\
\end{cases}
$$
And the inverse of $F(x)$ is
$$
F^{-1}(x) = 
\begin{cases}
\sqrt{2x} & 0\leq x \leq \frac{1}{2} \\
2-\sqrt{2-2x} & \frac{1}{2}<x\leq1\\
\end{cases}
$$

Draw i.i.d. samples from $U(0,1)$ and then plug into $F^{-1}(x)$ to get the desired sample. Codes and results:

```{r}
n <- 10
cdf <- function(x) {
  if (x >= 0 && x <= 0.5) {
    return(sqrt(2*x))
  }
  
  if (x > 0.5 && x <= 1) {
    return(2 - sqrt(2 - 2*x))
  }
  
}

sample <- cdf(runif(n))
```

The desired sample is
`r sample`