---
title: "Homework 5"
author: "Yiran Liu"
date: "3/9/2020"
output:
  pdf_document:
    latex_engine: xelatex
mainfont: Georgia
monofont: Sarasa Term CL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

We would take the state equation as the proposal function.

```{r, question1}
u0 <- 0
w0 <- 1
Y <- c(-0.4, -1.4, 1.1, -0.3, 0.7, -0.1, 1.3, 0.3, 0.1, 1.2, 2.0, 0.9, -0.4, 0.9, 1.0)
U <- matrix(NA, 10000, 15)
W <- matrix(NA, 10000, 15)
h <- rep(0, 15)
for (m in 1:10000){
  u <- c()
  w <- c()
  z <- rnorm(1)
  u[1] <- sample(c(z, rep(u0, 9)), 1)
  w[1] <- w0 * dnorm(Y[1], mean = u[1])
  for (i in 2:15){
    z <- rnorm(1)
    u[i] <- sample(c(z, rep(u[i - 1],  9)), 1)
    w[i] <- w[i-1] * dnorm(Y[i], mean = u[i])
  }
  U[m,] <- u
  W[m,] <- w
}
h <- apply(U*W,2,sum) / apply(W,2,sum)
h
```

## Question 2

* Naive Monte Carlo

```{r, question2}
x1 <- runif(1000)
est1 <- mean(1 / (1 + x1^3))
var1 <- var(1 / (1 + x1^3))
```

By the naive Monte Carlo, we get $\hat{\mu}$=`r est1` and its variance is `r var1`.

* Control variates method

```{r}
u <- runif(1000)
b <- cov(1 / (1 + u^3), 1 + u^3) / var(1 + u^3)
x2 <- runif(1000)
est2 <- mean(1 / (1 + x2^3) - b * ((1 + x2^3) - 5/4))
var2 <- var(1 / (1 + x2^3) - b * ((1 + x2^3) - 5/4))
```

By the control variate method, we get $\hat{\mu}$=`r est2` and its variance is `r var2`, which is significantly smaller than the variance in the naive method.

## Question 3

The cdf of Pareto distribution is
$$
F(x)=\int^{x}_{2}160t^{-6}dt=1-32x^{-5},\text{  }2\leq x <\infty,
$$
and the inverse of $F(x)$ is
$$
F^{-1}(t)={(\frac{32}{1-t})}^{\frac{1}{5}}.
$$

* Naive Monte Carlo

```{r, question3}
n <- 1000
u <- runif(n)
x <- (32 / (1 - u))^(1/5)
est1 <- mean(x)
var1 <- var(x) / n
```

By the naive Monte Carlo, we get $\hat{\mu}$=`r est1` and its variance is `r var1`.

* Antithetic variates

```{r}
u <- runif(n/2)
x1 <- (32 / (1 - u))^(1/5)
x2 <- (32 / u)^(1/5)
est2 <- mean((x1 + x2)/2)
var2 <- (var(x1) + var(x2) + 2*cov(x1, x2))/(4*n)
```

By the antithetic variate method, we get $\hat{\mu}$=`r est2` and its variance is `r var2`, which is significantly smaller than the variance in the naive method.