---
title: "Homework 3"
author: "Yiran Liu (netID: yiranl7)"
date: "2/10/2020"
output:
  pdf_document:
    latex_engine: xelatex
mainfont: Georgia
monofont: Sarasa Term CL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

In this question, $h(x) = \sin(\sqrt{x})$ and $\pi(x)=I_{[0,1]}(x)$. We intend to use $Beta(\frac{3}{2},1)$ as the proposal distribution. In comparison, we also apply the native Monte Carlo method to estimate $\mu$. Codes and results:

```{r}
h <- function(x) {
  sin(sqrt(x))
}

n <- 1000
## importance sampling
x1 <- rbeta(n, 1.5, 1)
w <- 1/dbeta(x1, 1.5, 1)
mu <- mean(w*h(x1))
## direct monte carlo
x2 <- runif(n)
mu2 <- mean(sin(sqrt(x2)))
se1 <- sd(w*h(x1))/sqrt(1000)
se2 <- sd(x2)/sqrt(1000)
```

The result shows that by applying importance sampling, the standard error reduced from `r se2` (by native Monte Carlo method) to `r se1`.

## Question 2

In HW2, the sample size to generate 1000 accepted samples was $N =1507$. For this question, we use normal distribution $N(0.68, 1)$ as the proposal distribution. Codes and results:

```{r}
l <- function(x) {
  1/(pi * (1 + x^2)) * 1/(sqrt(2 * pi))^5 * exp((-5 * (x - 0.68)^2 - 2.948)/2)
}
n <- 1507
x <- rnorm(n)
w <- l(x) / dnorm(x)
mu <- sum(w*x)/sum(w)
se <- sqrt((var(w * x) + mu^2 * var(w) - 2 * mu * cov(w, w*x)) / (n * (mean(w))^2))
```

The estimate of mean of $\pi(x)$ is `r mu` and the standard error is `r se`.

## Question 3

Use standard normal distribution as the proposal distribution. Codes and results:

```{r}
n <- 1000
x <- rnorm(n, 0, 1)
w <- exp(-0.5 * x^8) / dnorm(x, 0, 1)
mu <- sum(w*(x^2))/sum(w)
se <- sqrt((var(w * x^2) + mu^2 * var(w) - 2 * mu * cov(w, w*x^2)) / (n * (mean(w))^2))
```

The estimate is `r mu` and the standard error is `r se`.